---
title: "Washington County Map"
output:
  distill::distill_article:
    self_contained: TRUE
    toc: FALSE
    style: style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      error = FALSE,
                      cache = TRUE,
                      self_contained = TRUE,
                      fig.height = 6,
                      out.width = "100%")
```

```{r packages}
library(tidyverse)
library(tidycensus)
library(janitor)
library(scales)
library(sf)
library(tigris)
library(leaflet)
library(mapdeck)
```

# Race/Ethnicity

One map for each group based on previous spreadsheet

```{r, layout="l-screen-inset"}

mapdeck(style = mapdeck_style("light"),
        token = "pk.eyJ1IjoiZGdrZXllcyIsImEiOiJ2WGFJQ2U0In0.ftoZlfudaEIJL7OEf-Mw3Q",
        pitch = 25,
        location = c(-123.096819, 45.5),
        zoom = 9)
# %>% 
#   add_sf(data = non_white,
#          tooltip = "school",
#          fill_colour = "mdn_vl_",
#          fill_opacity = 150,
#          stroke_width = 50,
#          stroke_colour = "#ffffff",
#          auto_highlight = TRUE,
#          update_view = FALSE,
#          palette = "magma",
#          highlight_colour = "#ffffff99",
#          na_colour = "#fafafa",
#          legend = list(fill_colour = TRUE),
#          legend_options = list(title = "Non-White Students"))
```

# Poverty



## 100% FPL

One map for schools

## 200% FPL

Still do this map for individual schools even though not for composite score

# Single-Parent Families

Make map

# SNAP

Just SNAP

# Disability (EI/ESCE)

As table if just at district level
Mapped if at elementary school catchment area

# Homelessness

Evan is going to send it to me

# Composite Map

Calculate score with:

- Racial/ethnic diversity
- 100% FPL
- Single-parent families
- SNAP
- Rent burden



