---
title: "Washington County Map"
output:
  distill::distill_article:
    toc: FALSE
    style: style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      error = FALSE,
                      out.width = "100%",
                      fig.height = 6)
```

```{r packages}
library(tidyverse)
library(tidycensus)
library(janitor)
library(scales)
library(mapdeck)
library(devtools)

load_all()
```

# Introduction

I'll add more text that explains the maps here. Things like:

- Dots are schools
- Hover over any area to get more info
- Issues with data

# Race/Ethnicity

The following map shows the composite race/ethnicity measure for all groups. 

```{r, layout="l-page"}
race_ethnicity_composite %>% 
  make_map()
```


## Black


```{r, layout="l-page"}
race_ethnicity_black %>% 
  make_map()
```

## American Indian and Alaska Native

```{r, layout="l-page"}
race_ethnicity_aian %>% 
  make_map()
```

## Asian

```{r, layout="l-page"}
race_ethnicity_asian %>% 
  make_map()
```

## Caucasian

```{r, layout="l-page"}
race_ethnicity_white %>% 
  make_map()
```

## Hispanic

```{r, layout="l-page"}
race_ethnicity_latino %>% 
  make_map()
```

# Poverty



## 100% FPL

```{r, layout="l-page"}
children_in_poverty_100_fpl %>%
  make_map()
```

## 200% FPL

```{r, layout="l-page"}
children_in_poverty_200_fpl %>%
  make_map()
```


# Single-Parent Families

```{r, layout="l-page"}
single_parent %>%
  make_map()
```

# SNAP

```{r, layout="l-page"}
snap %>%
  make_map()
```

# Disability (EI/ESCE)

As table if just at district level
Mapped if at elementary school catchment area

```{r, layout="l-page"}
disability %>%
  make_map()
```

# Homelessness

```{r, layout="l-page"}
homeless %>%
  make_map()
```

# Composite Map

Calculate score with:

- Racial/ethnic diversity
- 100% FPL
- Single-parent families
- SNAP
- Rent burden


